---
title: "exercise-05"
author: "Jessica V"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

# Step 1
```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv"
d <- read_csv(f, col_names = TRUE)
head(d)

```

# Step 2


```{r}

a <- d%>%
  filter(startYear >1919) %>%
  filter(startYear <1980) %>%
  filter(runtimeMinutes <=180)%>%
  filter(runtimeMinutes >=60) %>%
  mutate(Decade = paste0(floor(ifelse(startYear<2000, startYear-1900, startYears)/10)*10, "'s")
  
  
```

Step 4

Use a one-line statement to calculate the 
population mean and population standard deviation
in runtimeMinutes for each decade and save the results in a new dataframe called results.

```{r}
#mean(a$runtimeMinutes)

aa <- a$runtimeMinutes[a$Decade == "20's"]
m <- mean(aa)
s <- length(aa)
error <- qnorm(0.975)*152/sqrt(n)
left <- m-error
right <- m+error

```




Step 3
plot histograms of the distribution of runtimeMinutes for each decade

```{r}
par(mfrow = c(1, 2))  # sets up two panels side by side
attach(d)  # lets us use variable names without specifying the data frame!
hist(runtimeMinutes) 

#freq = FALSE, col = "red", main = "Plot 1", xlab = "dec",
  #  ylab = "Run Time Minutes", ylim = c(0, 300))

#hist(log(startYear), freq = FALSE, col = "red", main = "Plot 2", xlab = "log(area)", ylab = "density",
  #  ylim = c(0, 100))
```
```{r}
attach(a)
#library(Hmisc)
hist.data.frame(a_startYear)
a_runtime <- a(runtimeMinutes)
```

```{r}
a %>% 
  hist(startYear) %>% 
  ggplot(aes(value)) + 
  geom_histogram() + 
  facet_wrap(~name, scales = "free")
```


```{r}
hist(log(population), freq = FALSE, col = "red", main = "Plot 1", xlab = "log(population size)",
    ylab = "density", ylim = c(0, 0.2))
hist(log(area), freq = FALSE, col = "red", main = "Plot 2", xlab = "log(area)", ylab = "density",
    ylim = c(0, 0.2))

```

```{r}
attach(a)
par(mfrow=c(1, 5))

# Create side-by-side histograms
hist(startYear$runtimeMinutes, main="Histogram of ", xlab="Decade", ylim = c(0, 300),
     ylab="runtimeMinutes", col="lightblue", xlim=c(10, 35))
#hist(mtcars$hp, main="Histogram of Horsepower", xlab="Horsepower", 
   #  ylab="Frequency", col="lightgreen")
```
```{r}
#par(mfrow = c(1, 2))  # sets up two panels side by side
attach(d)  # lets us use variable names without specifying the data frame!
#hist(runtimeMinutes, freq = FALSE, col = "red", main = "Plot 1", xlab = "Decade",
    #ylab = "runtimeMinutes", ylim = c(0, 300))
#hist(log(area), freq = FALSE, col = "red", main = "Plot 2", xlab = "log(area)", ylab = "density",
   # ylim = c(0, 0.2))

hist(runtimeMinutes), freq = FALSE, col = "red", main = "Plot 2", xlab = "(runtimeMinutes)", ylab = "Decade",  ylim = c(0, 0.2))
#abline(v = mean(log(area), na.rm = TRUE))
```
```{r}
#hist(d$runtimeMinutes)
# load library ggplot2
#library(ggplot2)
  
# set theme
theme_set(theme_bw(12))
 
# create x vector
xAxis <- rnorm(500)             
 
# create groups in variable using conditional 
# statements
group <- rep(1, 500)              
group[xAxis >  20] <- 2
group[xAxis > 30] <- 3
group[xAxis > 40] <- 4
group[xAxis > 50] <- 5
group[xAxis > 60] <- 6
 
# create sample data frame
sample_data <- data.frame(xAxis, group) 
  
ggplot(a, aes(x=xAxis, fill = as.factor(group)))+
   geom_histogram( color='#e9ecef', alpha=0.6, position='identi
```
```{r}
a %>% 
  pivot_longer(cols =11 ) %>% 
  ggplot(aes(value)) + 
  geom_histogram() + 
  facet_wrap(~name, scales = "free")

```

#Extra Notes for JEssica below:
Trial and Error

ssss<- d%>%
  
filter(grepl(pattern = "^[>=1920]", startYear, )) %>%
  
filter(grepl(pattern = "[<=1979]", startYear, )) %>%

filter(grepl(pattern = "[>=60]", runtimeMinutes, )) %>%

filter(grepl(pattern = "[<=180]", runtimeMinutes, ))


notes
s<-filter(d, grepl(pattern = "[1920-1979]", startYear))

ss<-filter(d, grepl(pattern = "[>60]", runtimeMinutes))

sss<- filter(d, grepl(pattern = "[<180]", runtimeMinutes))

